---
title: "Long-term Fertilization and Mowing at WRC:  Plant Alpha Diversity Analyses - Wetland Species Only"
author: "Michael W. McCoy, Ariane Peralta, and Carol Goodwillie"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  fig_caption: yes
  pdf_document: null
header-includes:
---
Narrative for the analysis of data from a long-term ecological experiment conducted by Dr. Carol Goodwillie at ECU's West Research Campus. The experiment has monitored the change in plant communities over time. The experimental design is a 2x2 factorial experiment with 2 levels of disturbance (mowed or unmowed) and 2 levels of nutrient addition (fertilized and unfertilized). The experiment was arranged in 8 spatial blocks with each containing the 4 treatment plots. Three fixed quadrats were created within each plot that were sampled annually. The spatial blocks were arranged in two rows that differed in proximity to a ditch which is suspectected to have induced systematic differences in the hydrology among the sets of blocks (see Figure 1).Blocks near the ditch are drier than blocks far from the ditch, which are wetter.

Okay, so now lets test the wetland obligate plants seperately from the generalist and upland species.  If the ditch effect is only a drying affect we should see a significant ditch effect on the wetland plants but no significant effect on the upland plants.

## Initial Setup
```{r Initial Setup, include=FALSE}
rm(list=ls())
setwd("~/GitHub/WRC_FertMowing/analyses")
```

```{r prep_wetland_dat,echo=FALSE,eval=FALSE}
###Ask Carol to update wetland status.csv
plant <- read.csv("../data/WRC_Importance.csv")
str(plant)

wetplant <- read.csv("../data/wetland status.csv")
head(wetplant)
str(wetplant)
obligate_1=which(wetplant[1,]==1)
obligate_2=which(wetplant[1,]==2)
obligate=sort(c(obligate_1))
names(wetplant)

head(wetplant[,obligate+1])
##abundance
#
wet_abun <- plant[,obligate+1]
str(wet_abun)
library(tidyverse)
abund=wet_abun %>% mutate(sumrow = rowSums(.),meanrow = rowMeans(.))


WRC_wetdat.2 <- data.frame(plant[,1:9],abundance=abund$sumrow, meanimp=abund$meanrow)
head(WRC_wetdat.2)
range(WRC_wetdat.2$meanimp)
range(WRC_wetdat.2$abundance)
str(WRC_wetdat.2)
write.csv(WRC_wetdat.2, file="../data/WRC_wetPlant_abundance.csv")
```

Now draw a picutre of the wetland plant data.

```{r , echo=FALSE}
ggplot(WRC_wetdat.2,aes(x=Year,y=meanimp,group=Ditch,colour=factor(Ditch)))+
  stat_summary(fun.data=mean_cl_boot,size=1)+facet_wrap(~Fertilizer*Mowing,labeller = label_both)
ggplot(WRC_wetdat.2,aes(x=Year,y=abundance,group=Ditch,colour=factor(Ditch)))+
  stat_summary(fun.data=mean_cl_boot,size=1)+facet_wrap(~Fertilizer*Mowing,labeller = label_both)
```

First we will look at diversity using richness as our measure of diversity...
```{r}
WRC_wetdat.2$Year_sc=scale(WRC_wetdat.2$Year, center = TRUE, scale = TRUE)
WRC_wetdat.2$BLOCK=as.factor(WRC_wetdat.2$BLOCK)
WRC_wetdat.2$Fertilizer=as.factor(WRC_wetdat.2$Fertilizer)
WRC_wetdat.2$Mowing=as.factor(WRC_wetdat.2$Mowing)
WRC_wetdat.2$Ditch=as.factor(WRC_wetdat.2$Ditch)
WRC_wetdat.2$QUADRAT..=as.factor(WRC_wetdat.2$QUADRAT..)
str(WRC_wetdat.2)
```

Wow these data are going to be tricky.  These are bound between zero and 1, and they appear to be zero inflated.  There also is no meaningful denominator associated with the proportions so questionable as to wehther we can assume binomial error.  So, I will attemtp to analyse them using a mixed model betaregression approach using glmmtmb.  THis is a brand new tool so...lets see what happens?

```{r, echo=FALSE}
#install.packages("TMB")
#install.packages("glmmTMB",type="source")
library(glmmTMB)
eps=1e-3 
squeeze=function(x){ (x*(length(x)-1)+0.5)/length(x)}
squeeze_2=function(x,eps){ (x+eps)/(1+2*eps)}
WRC_wetdat.2=WRC_wetdat.2 %>% mutate(meanimp_sq = squeeze(meanimp),meanimp_sq2=squeeze_2(meanimp,eps))
```

```{r}
wp1=glmmTMB(meanimp_sq~(Year_sc * Mowing * Fertilizer) + Ditch + (1|Ditch/BLOCK/QUADRAT..), data=WRC_wetdat.2,ziformula = ~1, family=beta_family(link = "logit"))
wp2=glmmTMB(meanimp_sq~(Year_sc * Mowing * Fertilizer) + Ditch + (1|Ditch/BLOCK/QUADRAT..), data=WRC_wetdat.2,ziformula = ~., family=beta_family(link = "logit"))
wp3=glmmTMB(meanimp_sq~(Year_sc * Mowing * Fertilizer) + Ditch + (1|Ditch/BLOCK/QUADRAT..), data=WRC_wetdat.2,ziformula = ~(Year_sc * Mowing * Fertilizer) + Ditch, family=beta_family(link = "logit"))
wp4=glmmTMB(meanimp_sq~(Year_sc * Mowing * Fertilizer) + Ditch + (1|Ditch/BLOCK/QUADRAT..), data=WRC_wetdat.2,ziformula = ~0, family=beta_family(link = "logit"))

c(glance(wp1)[3],glance(wp2)[3],glance(wp3)[3],glance(wp4)[3])
```

```{r}
library(car)
Anova(wp4,type="III")

ae <- allEffects(wp4)
ae
plot(ae)
summary(wp4)
```


Model generated plot (i.e. plot with random effect variaince included and with boot strap CIs)

```{r, echo=FALSE,warnings=FALSE,message=FALSE}
wp4_CI <- confint(wp4,method="Wald")

predframe=predict(wp1,type="response",newdat=WRC_wetdat.2,se.fit=T)
pframe=data.frame(WRC_wetdat.2,fit=predframe$fit,
                  upr=predframe$fit+2*predframe$se.fit,
                  lwr=predframe$fit-2*predframe$se.fit)

int=Effect(c("Year_sc","Mowing","Fertilizer"),wp4)
inter=summary(int)
WRC_wetdat.2$chYear=as.factor(WRC_wetdat.2$Year)
c(-1.5,-0.7,0.1,0.9,2)
modterms=expand.grid(Year=c(2004,2007,2010,2013,2017),Mowing=c(0,1),Fertilizer=c(0,1))
estimates=c(inter$effect[,,1][,1],inter$effect[,,1][,2],inter$effect[,,2][,1],inter$effect[,,2][,2])
lwr=c(inter$lower[,,1][,1],inter$lower[,,1][,2],inter$lower[,,2][,1],inter$lower[,,2][,2])
upr=c(inter$upper[,,1][,1],inter$upper[,,1][,2],inter$upper[,,2][,1],inter$upper[,,2][,2])
eff=data.frame(modterms,estimates,lwr,upr)

# New facet label names forvariable
fert.labs <- c("Unfertilized", "Fertilized")
names(fert.labs) <- c("0", "1")
mow.labs <- c("Unmowed", "Mowed")
names(mow.labs) <- c("0", "1")
p1=ggplot(data=eff,aes(x=Year,y=estimates,ymin=lwr,ymax=upr))+
  geom_smooth(method="lm",color="black")+
  facet_grid(Fertilizer~Mowing,labeller = labeller(Mowing = mow.labs, Fertilizer = fert.labs))+
  theme_bw()+ylab("Importance Score")+
 theme(axis.text.x = element_text(size=18,colour = "black", hjust=1,angle=45),
        axis.title.x = element_text(size=18,face="bold")) +
  theme(axis.text.y = element_text(size=18,colour = "black"),
        axis.title.y = element_text(size=18,face="bold")) +
  theme(strip.text.x = element_text(size = 18,face="bold"),strip.text.y = element_text(size = 18,face="bold"))+
   theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
        #panel.border = element_rect(size = 1)
     )+ggtitle("Time x Mowing x Fertilizer Effect")
p1

p3=ggplot(data=WRC_wetdat.2,aes(x=Year,y=meanimp,color=Ditch, shape=Ditch))+
  stat_sum(aes(size=factor(..n..)))+
  #geom_violin(position="identity")+
  facet_grid(Fertilizer~Mowing,labeller = labeller(Mowing = mow.labs, Fertilizer = fert.labs))+
   theme_bw()+ylab("Importance Score")+ theme(legend.position = "none") +
   scale_colour_manual(labels = c("wet_0ditch","dry_1ditch"), 
                    values = c("black", "gray")) +
    scale_shape_manual(labels = c("wet_0ditch","dry_1ditch"), 
                   values = c(21,17)) + 
 theme(axis.text.x = element_text(size=18,colour = "black", hjust=1,angle=45),
        axis.title.x = element_text(size=18,face="bold")) +
theme(axis.text.y = element_text(size=18,colour = "black"),
        axis.title.y = element_text(size=18,face="bold")) +
  theme(strip.text.x = element_text(size = 18,face="bold"),strip.text.y = element_text(size = 18,face="bold"))+
   theme(#plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(size = 1))+ggtitle("Raw Data")
p3
```

```{r}
d=Effect("Ditch",wp4)
dd=summary(d)
ditcheff=data.frame(Ditch=c("wet (0ditch)", "dry (1ditch)"),est=c(dd$effect[1],dd$effect[2]),lwr=c(dd$lower[1], dd$lower[2]),upr=c(dd$upper[1], dd$upper[2]))

p2=ggplot(ditcheff,aes(x=Ditch,y=est,ymin=lwr,ymax=upr,group=Ditch))+geom_segment(x=2,y=dd$effect[1],xend=1,yend=dd$effect[2],size=1)+geom_point(size=3)+theme_bw()+geom_linerange(size=.5)+theme_bw()+ylab("Importance Score")+
theme(axis.text.x = element_text(size=18,colour = "black", hjust=0.5),
        axis.title.x = element_text(size=18,face="bold")) +
  theme(axis.text.y = element_text(size=18,colour = "black"),
        axis.title.y = element_text(size=18,face="bold")) +
   theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(size = 1))+ggtitle("Ditch Effect")
p2
#library(gridExtra)
#suppressWarnings(grid.arrange(p1,p2,p3))
#pdf("~/Desktop/Wetland_plants.pdf")
#suppressWarnings(p1)
#suppressWarnings(p2)
#suppressWarnings(p3)
#dev.off()
```
# Combine graphs
```{r}
library("cowplot") 

# align all plots vertically
p3.l <-p3 + theme(legend.position="top", legend.box = "horizontal")
plots <- align_plots(p1, p2, p3.l, align = 'v', axis = 'l')

# put together bottom row and then everything
p4 <- plot_grid(plots[[1]],plots[[2]],plots[[3]], rel_heights = c(1.1,1,1.3), labels = c("A","B","C"), ncol=1, align='v')
p4        

ggsave("../figures/ObligateWetlandSpp_Diversity.pdf", plot=last_plot(), device=NULL, path=NULL, scale=1,  width=7, height=15, dpi=300, limitsize=TRUE)
```